An alternative to classic Peclet number blending is the usage of a discontinuity 
capturing operator (DCO), or in the low Mach context a nonlinear stabilization 
operator (NSO). In this method, an artifical viscosity is defined that is a function 
of the local residual and scaled computational gradients. Viable usages for the NSO 
can be advection/diffusion problems in addition to the aforementioned RTE VMS approach.

The formal finite element kernel for a NSO approach is as follows,
\begin{equation}
  \sum_e \int_\Omega \nu(\mathbf{R}) \frac{\partial w}{\partial x_i} g^{ij} \frac{\partial \phi} {\partial x_j} d\Omega,
\label{nsoFEMForm}
\end{equation}
where $\nu(\mathbf{R})$ is the artifical viscosity which is a function of the pde fine-scale 
residual and $g^{ij}$ is the covariant metric tensor).

For completeness, the covariant and contravarient metric tensor are given by,
\begin{equation}
  g^{ij} = \frac{\partial x_i} {\partial \xi_k}\frac{\partial x_j} {\partial \xi_k},
\label{coVariant}
\end{equation}

 and

\begin{equation}
  g_{ij} = \frac{\partial \xi_k} {\partial x_i} \frac{\partial \xi_k} {\partial x_j},
\label{coVariant}
\end{equation}

where $\xi = (\xi_1, \xi_2, \xi_3)^T$.
%
The form of $\nu(\mathbf{R})$ currently used is as follows,
\begin{equation}
  \nu = \sqrt{ \frac{\mathbf{R_k} \mathbf{R_k}} {\frac {\partial \phi}{\partial x_i} g^{ij} \frac{\partial \phi}{\partial x_j}} }.
\label{nuOne}
\end{equation}
The classic paper by Shakib (~\cite{Shakib:1991}) represents the genesis of this method which was done 
in the accoustically compressible context.

A residual for a classic advection/diffusion/source pde is simply the fine scale residual 
computed at the gauss point,
\begin{equation}
 \mathbf{\hat R} = \frac{\partial \rho \phi}{\partial t} + \frac{\partial}{\partial x_j} (\rho u_j \phi - \mu^{eff} \frac{\partial \phi}{\partial x_j}) -S
 \label{nsoResidual}
\end{equation}
Note that the above equation requires a second derivative whose source is the 
diffusion term. The first derivative is generally determined by using projected 
nodal gradients. As will be noted in the pressure stabilization section, the advection 
term carries the pressure stabilization terms. However, in the above equation, these 
terms are not explicity noted. Therefore, an option is to subtract the fine scale 
continuity equation to obtain the final general form of the source term,
\begin{equation}
 \mathbf{R} = \mathbf{\hat R} - \phi (\frac{\partial \rho}{\partial t} + \frac{\partial \rho u_j }{\partial x_j}).
 \label{nsoResidual}
\end{equation}

An alternative to the fine-scale PDE is a form that is found by differencing the linearized
form of the residual from the nonlinear residual,
\begin{equation}
 \mathbf{R} = \frac{\partial \rho u_j \phi }{\partial x_j} - (\phi \frac{\partial \rho u_j }{\partial x_j} + \rho u_j \frac{\partial \phi}{\partial x_j}).
 \label{nsoResidualAlt}
\end{equation}
The above resembles a commutation error in the nonlinear advection term.

In general, the NSO-$\nu$ is prone to percision issues when the gradients are very close to 
zero. As such, the value of $\nu$ is limited to a first-order like value. This parameter is 
proposed as follows: if an operator were written as a Galerkin (un-stabilized) plus a diffusion 
operator, what is the value of the diffusion coefficient such that first-order upwind is obtained 
for the combined operator? This upwind limited value of $\nu$ provides the highest value that 
this coefficient can (or should) be. The current form of the limited upwind $\nu$ is as follows,

\begin{equation}
  \nu^{upw} = C_{upw}(\rho u_i g_{ij} \rho u_j )^{\frac{1}{2}}
\label{nsoFVForm}
\end{equation}
where $C_{upw}$ is generally taked to be ~0.1.

Using a piecewise-constant test function suitable for CVFEM and EBVC schemes (the reader is refered to 
the VMS RTE section), Eq.~\ref{nsoFEMForm} can be written as,
\begin{equation}
  -\sum_e \int_\Gamma \nu(\mathbf{R}) g^{ij} \frac{\partial \phi} {\partial x_j} n_i dS.
\label{nsoFVForm}
\end{equation}
%
A fourth order form, which writes the stabilization as the difference between the Gauss-point gradient 
and the projected nodal gradient interpolated to the Gauss-point, is also supported,
\begin{equation}
  -\sum_e \int_\Gamma \nu(\mathbf{R}) g^{ij} (\frac{\partial \phi} {\partial x_j} - G_j \phi ) n_i dS.
\label{nsoFVForm4th}
\end{equation}

\subsection{NSO Based on Kinetic Energy Residual}
An alternative formulation explored is to share the general kernal form shown in Equation~\ref{nsoFVForm4th},
however, compute $\nu$ based on a fine-scale kinetic energy residual. In this formulation, the fine-scale
kinetic energy residual is obtained from the fine-scale momentum residual dotted with velocity. As with the 
continuity stabilization approach, the fine-scale momentum residual is provided by Equation~\ref{fineScaleMomentum}.
Therefore, the fine-scale kinetic energy is written as,
\begin{equation}
 \mathbf{R}_{ke} = \frac{u_j(\frac{\partial p} {\partial x_j} - G_j p )}{2},
\label{fineScaleKe}
\end{equation}
while the denominator for $\nu$ now includes the gradient in ke,
\begin{equation}
  \nu = \sqrt{ \frac{\mathbf{R}_{ke} \mathbf{R}_{ke}} {\frac {\partial ke}{\partial x_i} g^{ij} \frac{\partial ke}{\partial x_j}} }.
\label{nuKe}
\end{equation}
The kinetic energy is simply given by,
\begin{equation}
  ke = \frac{u_k u_k}{2}
\label{keForm}
\end{equation}

The kinetic energy form of $\nu$ is used for all equation sets with transformation by usage of a 
turbulent Schmidt/Prandtl number.

\subsection{Local or Projected NSO Diffusive Flux Coefficient}
While the NSO kernel is certainly evaluated at the subcontrol surfaces, the evaluation of $\nu$ can be 
computed by a multitude of approaches. For example, the artificial diffusive flux coefficient 
can be computed locally (with local residuals and local metric tensors) or can be projected 
to the nodes (via an $L_{oo}$ or $L_2$ projection) and re-interpolated to the gauss points. 
The former results in a sharper local value while the later results in a more filtered-like value.
The code base only supports a local NSO $\nu$ calculation.

\subsection{General Findings}
In general, the NSO approach seems to work best when running the fourth-order option as the second-order 
implementation still looks more diffuse. When compared to the standard MUSCL-limited scheme, the NSO
is the preferred choice. More work is underway to improve stabilization methods.
Note that a limited set of NSOs are activated in the code base with specific
interest on scalar transport, e.g, momentum, mixture fraction and static enthalpy transport. 
When using the $4^{th}$  order method, the consistent mass matrix approach for the projected 
nodal gradients is supported for higher order.
